\documentclass[oneside]{book}
\usepackage{graphicx}

\begin{document}

{
\large
The Train Janitor.  By Samuel Alexander
\par
}

\vspace{2mm}
As I stepped on the platform, I bumped into a slavelike janitor and saw he was listening to
my mp3 player.  I'd lost it the week before, when I rode into this place, I could tell it by the distinct
logo, an apple with a bite out.  As for the janitor, he'd fit in a concentration
camp.  He wore rags, he pushed a rented broom, too poor to afford a broom of his own, he was compelled to
give up the lion's share of his wages on rent.  There was a light of
intelligence in his eyes, and an aura about him like he'd seen better days.

``I say, good fellow, thanks for finding my mp3 player,'' I grabbed the device, wresting it
from his cold, weak hands.  ``Assuming you really did `find' it, I mean.
For all I know you probably stole it.''

``I found that player at this station four days ago,'' he said.  ``I'm glad you've come to reclaim it, I
checked Lost and Found every day in case its owner dropped by.''

``Yeah, well, I guess I owe you one,'' I reached into my shirt pocket and produced a breath mint, I tossed it
at the man and he caught it in his mouth, accustomed to performing for whatever charity he could get on the tracks.

``What made you decide to be a janitor here?'' I asked him offhanded, there was some time and I didn't
recognize the tracks on the mp3 player.

``Decide?'' said the man.  A passerby flicked a cigarette in his face, he caught it and put it in a little pouch
of cigarette butts, probably some kind of barter token on whatever street he slept on.  ``I
did not `decide', sir.  And heed my words, let them be a warning to you.  They pressed me into this miserable
position when they caught me riding first class with a second class ticket.  Before that I was a promising
student, I was making straight A's in a school you've heard of, I had a glamorous
internship.  I begged them to spare me, let me off with a warning.  Don't you forget, sir, in this country they
don't take kindly to cheaters.  They lashed me fourty times and then sentenced me to a life of servitude.
Sir, they enslaved me, and my extended family too.  My poor parents, bless them, torn from the golden years they'd
built over decades, now they polish train tracks.  My wife, the way she still dotes over me, scrubbing my rags on
days when the trains don't run.  She used to be beautiful sir, and I'll swear before any man, she's still got a
spark in her eyes, like the first day I met her, dancing and swinging her hips at the Cactus Club Grill.  Our
daughter,'' here he couldn't hold back his tears, they dripped onto the platform, adding to the cleaning he'd have
to do today, ``our daughter was just a babe, she couldn't understand as they barked orders at her, she
didn't have the vocabulary yet.  She crawled after us, washrag in her teeth, but she didn't
last long.  The rigours of work wore her all away, now she polishes God's trains in Heaven.''

I handed the man a McDonalds napkin to wipe his tears, it was embarrassing seeing the shameless display.  I looked
around self-consciously to check no-one from the hotel was watching, I wouldn't want to be associated
with this blubbering `what measure is a janitor'.  With a great sob he tore my napkin, as if it represented all the endless strawmen
he blamed for his lousy life's lot.  The pieces fluttered to the platform, did he not realize he was only
adding to his work?  ``Get yourself together, man,'' I said, ``remember where you are!''

He sniffed hard, wiped his nose on his sleeve.  ``Promise me,'' he said, ``promise me you won't repeat
my mistake.  Respect the tickets!  Don't sneak into first class.  Be content with second, and smile and
bow when the conductor comes by.  Sir, my own life is forfeit, don't let it be in vain.  Promise me, I beg you,
promise me you won't cheat the rail.''

``I promise,'' I said, stepping toward the yellow line, the PA was announcing my train.

``No!'' said the janitor grabbing my arm, ``That's not enough!  Sir, you don't know the gravity of the situation.
Sir, I can see you're an important man.  They'd let you cut in line at the Cactus Club Grill.  But sir,
do you know you could lose it in a day?  They don't give you first warnings here, sir.  One step in
first class, and they whisk it all away.  They have extradition treaties sir (I can tell you're not from
here, don't think it'll protect your family).  They have long memories, and they always keep grudges.  So
promise, sir, on your mother's grave.  Promise by all that is holy, promise by all that is good and wholesome in
this world, by the charitable hand, by the spark of hope that unites all our hearts, swear to me sir, that you will
not cheat the railway!''

``I swear by my mother's grave,'' I said.  ``I swear by all the virtues of man, I swear on all that is solemn
and dignified, I promise with my right hand on every page of the Bible and the apocrypha too, beneath the eyes of
the saints and the prophets, my word and contract: I will not cheat the railway.''

To prove my sincerity I took my pocket knife, I opened the blade, and with a bit of effort I nicked
my thumb.  I pressed the palm of the thumb to the man's sweaty brow.  ``This blood is witness,'' I said, ``to my
solemn promise.  I will not sit in first class with a second-class ticket.''

Then the train arrived.  I stepped in, glad to be out of that grave situation.  I walked down one car, so the
janitor could no longer see me, then I ascended the steps to the first-class cabin.  Comrades in fancy suits
smiled at me in recognition.  I spotted two business partners who'd come for the same conference as me,
one of them had attended my talk.  Beside them, (the blood rushed to my $\ldots$cheeks!), a young lady I had more
than casual acquaintances with from the Hilton Executive.  If I'd known she'd be here, I'd have worn my
triple-breasted suit, not this shabby two-breasted thing.  ``Say, friend, a gin and tonic and one for the lady,''
I waved down the waiter, I pressed a stunning wad of cash in his palm, ``Keep the change,
appreciation for the fine work you do.''  In the corner of my eye I caught the dame's look of approval,
I felt like I was walking on air.

``Come sit with us!'' said the man from my talk.  ``Do you guys know him?
You should've seen his talk!  We're looking at the future of the world here!''

``Now, now,'' I said, sitting down humbly.  ``The world?  You're flattering me probably.  I'll be happy
to be the future of the north hemisphere.''

There was a grate on the floor, through which first-class passengers could look down on second
and jeer at them or throw peanuts for sport.  My eye drifted and my heart suddenly
froze.  Looking back up at me I saw none other than the janitor.  So his shift was over, eh?  And he had to
choose exactly that seat!  Curse God in heaven, and all the saints and prophets!

``That was a cruel joke, sir,'' the janitor said.  ``You could have just told me you had a first class
ticket!''  ``Silence!'' said a sharp voice.  The oarmaster stepped into sight underneath us, he gave the janitor
a good lashing.  Defeated, the janitor looked at me in silence after that.  The PA announced our departure,
the janitor grabbed his oar and started to row.  That's how they save money here, the second classers row like
it's a Roman galleon.

As country rushed past, I regaled my comrades with tales of adventure.  They ``oohed'' and ``ahhed'' at
every word, and my lady friend placed her soft hand in mine.  Across from us, a spritely pair of
Italian tourists cocked their heads to understand my English, talking to each other excitedly whenever they made
out a glamorous detail.  They cheered me like a hero-celebrity, but I couldn't enjoy it proper,
my eye kept meeting the janitor's.

Then my heart jumped into my throat when I saw the conductor.  He was coming through on a ticket inspection.
My friends laughed lighthearted when I stammered:  ``I must have left my ticket
at the station or something.''

``What are you joshing about,'' laughed my talk-attender, with a friendly jab in the ribs.  ``Don't mind him,
gov'ner, when you're as big as he is, you forget little things!''  He grabbed my wallet
and pulled out the ticket.

But where all expected a flash of gold, instead came a strip like crude toilet paper.  It was my second-class
ticket.  I saw Hilton Girl's eyes, wide with shame.  The Italians turned to each other with a torrent of
``Mama Luigi!''s.  My business friends both stood at once, ``I've got a call to make,'' they said in one voice,
a beeline for the next car.  Second class passengers were pointing up at me and laughing, hocking
black loogies up at the grate, hooping and hollering like I was some circus freak.  Even the bumpkins outside
the train laughed, pointing at me through the train window.  Only the janitor
sat still, tears streaming from those weathered eyes still locked on me.

``Infidel!  Do you know what you've done?'' said the conductor.  ``Men like you are the root of all bad!
Do you see how the steerage riders struggle at the oars?  You can watch the life drain
from `em!  All to subsidize a leech like yourself.  Why, just two minutes ago, an
old man died of exhaustion.  He's left five orphans, they'll
starve and it's your fault!  Do you know what we do here, to people who cheat?''

I threw myself at the conductor's feet, the side of my face pressed against the grate as I slobbered on his shiny
steel-toed boots.  Hilton Girl went to his side, she leaned in real close and intimate-like and whispered
something in his ear, from the look on her face she was not on my side.  With a `Hmmph!' she
lifted her elegant rose-red heel, and she drove it into my grovelling ass.  The passengers below gave a raucous
cheer, and she flashed them a ruby smile.

``Please, sir,'' I said, genuinely sobbing.  ``Give me another chance.  I don't want to be a janitor.''  My tears dripped
down on the cheeks of the janitor, mingling with his own.  ``Take my family,
but don't take me!''  I hugged the conductor's leg, I french-kissed his boot, my tongue squirming between the heel and the sharp
metal grate.  ``Sir, I'm a wealthy man.  I only bought that ticket from miserly vice, sir.  I know my actions are
inexcusable.  I know if you judge me by them, I'm a janitor's janitor.  But I didn't do it out of poverty.
Sir, I could have easily paid first class for everyone on this train, and every other train running today.''

``I see,'' said the conductor.  ``Then you really are a first class man.''  As if doubting me, he scooped up my wallet,
whistled as he leafed through the bills.  ``Yes, I see you're really first class.  You are innocent of cheating the railway.
Never-the-less, you have mocked our ceremonies.  You should have purchased the right ticket.  Your behavior
is beyond the pale.  If you were a second-classer, I'd nail you to the axle.  Still I can't just leave you unpunished.
I sentence you to spend the rest of this ride down in second.''

I wailed like the words cut through me.  ``Not that, sir!''  I clutched his pantleg like a naughty infant son.  ``Sir, send
her instead,'' I indicated my former girlfriend, ``but please sir, I beg you, don't make me sit there!''

The conductor crouched down and grabbed my face, he turned it toward an old woman in the second class.  She
was a grandmother type, it was heart-wrenching to see her labor at her oar.  From the lines on her face you'd
think she was twice her true age, which was eighty.  She labored one-handed, with her other arm she cradled an emaciated
baby, the baby kept hunting for a nipple, every now and then the oar would slip and hit the baby on the head,
inspiring a hungry cry.  ``Do you see that woman?'' said the conductor.  ``That woman is my mother.  Do you know
you've insulted her?  You've told me the place where my own mother sits isn't good enough for you.  I'm a man,
can I just take that lightly?''

But the old woman gave us some signal.  The conductor caught his tongue, he was silent a while.
``You're lucky,'' he said.  ``My mother has saved you.  I always said she was too good for her own good.
She doesn't believe in rudeness to guests, and you are a foreigner.  So this time your goose is saved.
But damn all the heavens if I catch you again.  You'd better swear up and down, you'd better swear
low and high, you'll never cheat the railway again.''

``Sir,'' I said, ``I promise by the shroud of Christ himself, I swear by the Archangel Gabriel, I give you my word and
even the copyright on it:  I will not cheat the railway again.''

With a grunt, the conductor let me go.  Muttering under his breath, he moved toward the stairs to check second-class
tickets.  I sat back and prepared for a show.  See, there's something I've left out in all this.  The truth
is I didn't pay for the ticket you saw.  I picked it from the janitor's pocket!

\end{document}
